Contextuality and the Born rule in fuzzy quantum theories

### 2017-03-22 prop_letter.tex
Some updates with Prof. Ortiz...

\documentclass[reprint, aps, prl,superscriptaddress, showpacs,
showkeys]{revtex4-1} 
\usepackage{fullpage}
\usepackage[T1,T2A]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{mathrsfs}
\usepackage{bbold}
\usepackage{wesa}
\usepackage{graphicx}
\usepackage{verbatim}
\usepackage{hyperref}

@@

In conventional quantum theory, when we measure a system described
by a density matrix~$\rho$ with respect to an observable~$\mathbf{O}$,
$\rho$ and the eigenspace of a eigenvalue~$\lambda$ completely
determines the probability of getting $\lambda$, which is computed
by the Born rule~$\mu_{\rho}^{\mathrm{\mathrm{B}}}\left(P_{\lambda}\right)=\Tr\left(\rho P_{\lambda}\right)$,
where $P_{\lambda}$ denotes the projector to the eigenspace of $\lambda$.
In another word, any density matrix~$\rho$ defines a quantum probability
measure~$\mu_{\rho}^{\mathrm{\mathrm{B}}}:\events\rightarrow[0,1]$,
where $\events$ is the set of all projectors in a Hilbert space.
Conversely, given any quantum probability measure~$\mu:\events\rightarrow[0,1]$
which should satisfy:
\begin{itemize}
\item $\mu(\mathbb{0})=0$. 
\item $\mu(\mathbb{1})=1$. 
\item For any projection $P$, $\mu\left(\mathbb{1}-P\right)=1-\mu\left(P\right)$. 
\item For a set of mutually orthogonal projections $\left\{ P_{i}\right\} _{i=1}^{N}$,
we have 
\begin{equation}
\mu\left(\sum_{i=1}^{N}P_{i}\right)=\sum_{i=1}^{N}\mu\left(P_{i}\right)\textrm{ .}\label{eq:QuantumProbability-Addition}
\end{equation}
\end{itemize}
Gleason's theorem asserts that there exists a unique mixed state~$\rho$
such that $\mu=\mu_{\rho}^{\mathrm{\mathrm{B}}}$ in a Hilbert space
of dimension $d\geq3$.

@@

Although considering IVPM is new for quantum, people have already
defined classical IVPMs~\cite{JamisonLodwick2004}. We find if we
restrict on a mutually commuting sub-event space~$\events'\subseteq\events$,
quantum IVPMs are one-to-one corresponding to the classical ones.
Despite restricting on the domain of quantum IVPMs, if we pick the
range as a limiting case, the collection of all singletons~$\mathscr{I}_{\infty}=\set{\left[x,x\right]}{x\in\left[0,1\right]}$,
then quantum IVPMs becomes real-valued probability measures which
case Gleason's theorem applies. In the next paragraph, we will show
the Kochen-Specker theorem is another limit case when the collection
of intervals are $\mathscr{I}_{\mathrm{D}}=\left\{ \imposs,\necess\right\} $,
where $\imposs=\left[0,0\right]$ and $\necess=\left[1,1\right]$.\newpage{}

@@

\begin{example}[Four-dimensional quantum three-interval-valued probability
measure]\label{ex:four-dimensional-three-value}Consider the orthonormal
basis~$\left\{ \ket{0},\ket{1},\ket{2},\ket{3}\right\} $ as defined
in table~\ref{table:commonBasis}. Let $\ket{\ps}=\left(\ket{0}+\ket{1}\right)/\sqrt{2}$,
$\ket{\ps'}=\left(\ket{0}+\ket{2}\right)/\sqrt{2}$. Then, we can
define a quantum IVPM~$\bar{\mu}:\events_{24}\rightarrow\mathscr{I}_{1}$
as follow. 
\begin{subequations}
\begin{eqnarray}
& & \bar{\mu}(\mathbb{0})=\bar{\mu}(\proj{0})=\bar{\mu}(\proj{\ps})=\bar{\mu}(\proj{\ps'})=\imposs\textrm{ ,}\\ 
& & \bar{\mu}(\mathbb{1})=\bar{\mu}(\mathbb{1}-\proj{0})=\bar{\mu}(\mathbb{1}-\proj{\ps})=\bar{\mu}(\mathbb{1}-\proj{\ps'})=\necess\textrm{ ,}\\ 
& & \bar{\mu}(P) = \unknown\qquad\textrm{otherwise.} 
\end{eqnarray}
\end{subequations}
To verify $\bar{\mu}$ is a quantum IVPM, first notice that $\bar{\mu}\left(P_{0}+P_{1}\right)\subseteq\bar{\mu}\left(P_{0}\right)+\bar{\mu}\left(P_{1}\right)$
implies equation (\ref{eq:QuantumInterval-valuedProbability-Inclusion})
by induction. Second, equation (\ref{eq:QuantumInterval-valuedProbability-Complement})
implies $\bar{\mu}\left(\mathbb{1}\right)\subseteq\bar{\mu}\left(P\right)+\bar{\mu}\left(\mathbb{1}-P\right)$.
Therefore, to prove equation (\ref{eq:QuantumInterval-valuedProbability-Inclusion}),
it is sufficient to check
\begin{subequations}
\begin{eqnarray}
& & \bar{\mu}\left(\proj{\psi_{0}}+\proj{\psi_{1}}\right)\subseteq\bar{\mu}\left(\proj{\psi_{0}}\right)+\bar{\mu}\left(\proj{\psi_{1}}\right)\textrm{ ,}\\
& & \bar{\mu}\left(\proj{\psi_{0}}+\proj{\psi_{1}}+\proj{\psi_{2}}\right)\subseteq\bar{\mu}\left(\proj{\psi_{0}}\right)+\bar{\mu}\left(\proj{\psi_{1}}+\proj{\psi_{2}}\right)
\end{eqnarray}
\end{subequations}
for orthogonal $\ket{\psi_{0}}$, $\ket{\psi_{1}}$, and $\ket{\psi_{2}}$.
This equation always holds because $\unknown$ is a subset of the
left-hand side of both of the equations.

>>>>

\documentclass[english,reprint, aps, prl,superscriptaddress, showpacs,
showkeys]{revtex4-1} 
\usepackage{fullpage}
\usepackage[T1,T2A]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{babel}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{mathrsfs}
\usepackage{bbold}
\usepackage{wesa}
\usepackage{graphicx}
\usepackage{verbatim}
\usepackage{hyperref}

@@

In conventional quantum theory, when we measure a property $\mathbf{O}$ of a 
physical system whose state is represented by a density matrix~$\rho$, the 
probability to obtain the value $\lambda$ is determined by the Born rule 
$\mu_{\rho}^{\mathrm{\mathrm{B}}}\left(P_{\lambda}\right)=\Tr\left(\rho P_{\lambda}\right)$,
where $P_{\lambda}$ denotes the projector onto the eigenspace corresponding 
to the eigenvalue $\lambda$ of the Hermitian operator $\mathbf{O}$~\citep{Born1983,peres1995quantum,544199,Jaeger2007}.

In other words, any density matrix~$\rho$ defines a quantum probability
measure~$\mu_{\rho}^{\mathrm{\mathrm{B}}}:\events\rightarrow[0,1]$,
where $\events$ is the set of all projectors in a Hilbert space~\citep{10.2307/2308516,gleason1957,Redhead1987-REDINA,Maassen2010}.
Conversely, given any quantum probability measure~$\mu:\events\rightarrow[0,1]$
which should satisfy:
\begin{itemize}
\item $\mu(\mathbb{0})=0$. 
\item $\mu(\mathbb{1})=1$. 
\item For any projection $P$, $\mu\left(\mathbb{1}-P\right)=1-\mu\left(P\right)$. 
\item For a set of mutually orthogonal projections $\left\{ P_{i}\right\} _{i=1}^{N}$,
we have 
\begin{equation}
\mu\left(\sum_{i=1}^{N}P_{i}\right)=\sum_{i=1}^{N}\mu\left(P_{i}\right)\textrm{ .}\label{eq:QuantumProbability-Addition}
\end{equation}
\end{itemize}
Gleason's theorem asserts that there exists a unique mixed state~$\rho$
such that $\mu=\mu_{\rho}^{\mathrm{\mathrm{B}}}$ in a Hilbert space
of dimension $d\geq3$~\citep{gleason1957,Redhead1987-REDINA,peres1995quantum}.

@@

Although considering IVPM is new for quantum, people have already
defined classical IVPMs~\cite{JamisonLodwick2004}. We find if we
restrict on a mutually commuting sub-event space~$\events'\subseteq\events$,
quantum IVPMs are one-to-one corresponding to the classical ones.
Despite restricting on the domain of quantum IVPMs, if we pick the
range as a limiting case, the collection of all singletons~$\mathscr{I}_{\infty}=\set{\left[x,x\right]}{x\in\left[0,1\right]}$,
then quantum IVPMs becomes real-valued probability measures which
case Gleason's theorem applies. In the next paragraph, we will show
the Kochen-Specker theorem~\citep{kochenspecker1967,peres1995quantum,Redhead1987-REDINA} is another limit case when the collection
of intervals are $\mathscr{I}_{\mathrm{D}}=\left\{ \imposs,\necess\right\} $,
where $\imposs=\left[0,0\right]$ and $\necess=\left[1,1\right]$.\newpage{}

@@

\begin{example}[Four-dimensional quantum three-interval-valued probability
measure]\label{ex:four-dimensional-three-value}Consider the orthonormal
basis~$\left\{ \ket{0},\ket{1},\ket{2},\ket{3}\right\} $ as defined
in table~\ref{table:commonBasis}. Let $\ket{\ps}=\left(\ket{0}+\ket{1}\right)/\sqrt{2}$,
$\ket{\ps'}=\left(\ket{0}+\ket{2}\right)/\sqrt{2}$. Then, we can
define a quantum IVPM~$\bar{\mu}:\events_{24}\rightarrow\mathscr{I}_{1}$
as follow. 
\begin{subequations}
\begin{eqnarray}
&   & \bar{\mu}(\mathbb{0})=\bar{\mu}(\proj{0})=\bar{\mu}(\proj{\ps}) \nonumber \\*
& = & \bar{\mu}(\proj{\ps'})=\imposs\textrm{ ,}\\ 
&   & \bar{\mu}(\mathbb{1})=\bar{\mu}(\mathbb{1}-\proj{0})=\bar{\mu}(\mathbb{1}-\proj{\ps}) \nonumber \\*
& = & \bar{\mu}(\mathbb{1}-\proj{\ps'})=\necess\textrm{ ,}\\ 
& & \bar{\mu}(P) = \unknown\qquad\textrm{otherwise.} 
\end{eqnarray}
\end{subequations}
To verify $\bar{\mu}$ is a quantum IVPM, first notice that $\bar{\mu}\left(P_{0}+P_{1}\right)\subseteq\bar{\mu}\left(P_{0}\right)+\bar{\mu}\left(P_{1}\right)$
implies equation (\ref{eq:QuantumInterval-valuedProbability-Inclusion})
by induction. Second, equation (\ref{eq:QuantumInterval-valuedProbability-Complement})
implies $\bar{\mu}\left(\mathbb{1}\right)\subseteq\bar{\mu}\left(P\right)+\bar{\mu}\left(\mathbb{1}-P\right)$.
Therefore, to prove equation (\ref{eq:QuantumInterval-valuedProbability-Inclusion}),
it is sufficient to check
\begin{subequations}
\begin{eqnarray}
&           & \bar{\mu}\left(\proj{\psi_{0}}+\proj{\psi_{1}}\right) \nonumber \\*
& \subseteq & \bar{\mu}\left(\proj{\psi_{0}}\right)+\bar{\mu}\left(\proj{\psi_{1}}\right)\textrm{ ,}\\
&           & \bar{\mu}\left(\proj{\psi_{0}}+\proj{\psi_{1}}+\proj{\psi_{2}}\right) \nonumber \\*
& \subseteq & \bar{\mu}\left(\proj{\psi_{0}}\right)+\bar{\mu}\left(\proj{\psi_{1}}+\proj{\psi_{2}}\right)
\end{eqnarray}
\end{subequations}
for orthogonal $\ket{\psi_{0}}$, $\ket{\psi_{1}}$, and $\ket{\psi_{2}}$.
This equation always holds because $\unknown$ is a subset of the
left-hand side of both of the equations.
